---
title: "penguin_analysis"
author: "Candela Ferrer"
date: "2024-10-16"
output: html_document
---

## Library loading
  When you click the **Knit** button a document will be generated that includes   both content as well as the output of any embedded R code chunks within the     document
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(palmerpenguins)
library(janitor)
library(here)

```

## Penguin data 
  here() is a very robust,reproducible and data-safe version of accessing wd. 
```{r penguins_data, echo=FALSE}
# To write data to csv
write.csv(penguins_raw, here("data", "penguins_raw.csv"))

#Load data 
penguins_raw <- read.csv(here("data","penguins_raw.csv"))

# Looking at data
head(penguins_raw) #first 6 rows
colnames(penguins_raw) # show col names

```


# Tidying-up data - common functions

```{r tidying_data, echo=FALSE}

# Removing columns

#penguins_noComments <- select(penguins_raw, -Comments) # removes comments column
#penguins_onlyComments <- select(penguins_raw, Comments) # removes everything but comments
#penguins_onlyComments <- select(penguins_raw, -starts_with("Delta")) # removes everything but comments

```

The problem with this is that it clogs up our environment with loads of different objects 

Instead, we can use the piping function...

#Piping 
%>% is the piping function. 
It means "and then"

```{r piping, echo=FALSE}
# Removing columns with pipe function: 
penguins_clean <- penguins_raw %>%
  select(-Comments) %>%
  select(-starts_with("delta")) 

```

# Clean names 
This unifies all the column names 
types of computer-readable names: 
  camelCase
  snake_case
  PascalCase
  
```{r tidy_col_names}
# Cleaning column names:
penguins_clean <- penguins_raw %>%
  select(-Comments) %>%
  select(-starts_with("delta")) %>%
  clean_names() #unifies col names

# Save clean data - CODE
write.csv(penguins_clean, here("data", "penguins_clean.csv"))
```

#Functions 
We can make functions in R

```{r Clean_data_function, echo=FALSE}
# Making a function to clean data

cleaning_penguin_columns <- function(raw_data){ # making new function
  raw_data %>% # generic - any raw data goes here
    clean_names() %>% 
    select(-comments) %>% 
    select(-starts_with("delta"))
}

penguins_clean <-cleaning_penguin_columns(penguins_raw)

colnames(penguins_clean)

```

# Using functions from function source 
It is always advisable to store functions in a separate file and then loading them into your script. 
This ensures that the functions do not get damaged 
```{r }
# Using our functions from the functions script

#creating a file 
#file.create(here("functions", "cleaning.R")) 
#sourcing file
source(here("functions", "cleaning.R"), echo = TRUE)

```

#Installing libraries in a reproducible way 
This keeps track of the packages being used in the session by creating a 'Project Library' separate from the System Library 
```{r renv, echo=FALSE}
#Install renv 
install.packages("renv")

renv::init()  # initializing renv
renv::snapshot()# Create a snapshot of the libraries in use => in rev.lock file 
renv::diagnostics() # run diagnostics
renv::restore() # Restore the snapshot to install the libraries

```

#Making new script 

```{r create_file, echo=FALSE}
# Creating new file
# file.create(here("functions", "cleaning.R"))

```


